---
title: "DA2-Assignment-2"
author: "Gyongyver Kamenar (2103380)"
geometry: "left=2cm,right=2cm,top=1cm,bottom=2cm"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE
)
```

```{r TASK, include=FALSE}
# Use the hotels-europe dataset and pick a city. Use hotel user rating to create a binary variable: highly_rated=1 if rating ≥ 4, 0 otherwise. Examine how high rating is related to the other hotel features in the data. 
# 
# Estimate linear probability, logit, and probit models with distance and stars as explanatory variables. You may add other variables if you wish. 
# Compare coefficients, marginal differences, and predicted probabilities, and discuss your results. 
# 
# This is an assignment that you may do alone or in a pair within your group. You shall write a short (max 1 page, excluding exhibits) report that describes your major decisions, your estimated models, interpretations and summary. You may have a descriptive data table (optional), a regression table and up to two graphs. All codes should be pushed to your version control repo, with appropriate commit policy.
# You need upload your report in pdf format. Add the link of your repo to the online text field.
```

```{r Libraries, include=FALSE}

# Loading packages 
if (!require("pacman")) {
  install.packages("pacman","PerformanceAnalytics")
}
library(pacman)
pacman::p_load(tidyverse, ggplot2, modelsummary,
               kableExtra, stargazer, knitr, fixest, corrplot,PerformanceAnalytics, pscl, mfx, ggpubr, tinytex, lspline)

```
#### Introduction
In this assignment I analyzed the relationship between hotel rating, distance from the city centre and stars of the hotel in Rome. 
The used dataset with hotels in Europe was provided by the textbook and the hotels in Rome were selected. I built 4 different models to predict the probability that the hotel is highly rated (at least 4) based on the distance and the number of stars. 

#### Data exploration
The datatable contains accomodations in Europe, so I filtered for  Rome and then I explored the accomodation types. I found several different types of accomodation in the datatset but I decided to narrow down the research question just to hotels. the analysis will better represent the hotels in Rome. Besides, I omitted observation with missing distance, rating or stars and also few observations with not integer stars.

To explore the data I checked the descriptive statistics (Table 1) of the most relevant numerical variables and also some dummy variables like offer, holiday, weekend and the highly rated, which is our variable of interest.
Furthermore, I examined the correlation between variables as illustrated in the first plot. The variable of my interest (highly_rated) is obviously highly correlated with rating, but also with stars and price, and there is significant but negative correlation with distance too. Finally, I plotted rating with distance and stars with the loess method, which suggest that splines on distance might be useful.

#### Models and interpretation
The first model is the linear probability model with distance and and stars as explanatory variables and highly rated as the dependent variable. The huge drawback of this model is that the estimeted probability is not limited so practically it can be below zero or above 1 as illustrated on Figure 2. The estimated coefficients are presented in Table 2. Both explanatory variables are significant at 1% level. The coefficient of distance means, that with the same stars hotels 1 unit further from the centre  are 0.036 less likely to be highly rated on average. Considering the stars coeffiecient,among hotels with the same distance, 1 more star hotels are 0.16 more likely to be highly rated on average.

The second model used the same variables but it is a logit model. Because of the functional formula the predicted probability is limited between 0 and 1. I calculated the average marginal differences (also included in Table 2) which can be interpreted similarly to the LPM. Based on the logit model, on average hotels 1 unit further are 0.045 less likely and hotels with 1 more start are 0.18 more likely to have high rating ceteris paribus. 

The third one is a probit model which is very similar to the logit. The functional form is also limits the values at 0 and 1, the estimated average marginal differences for distance and starts are -0.040 and 0.176 which are between the LPM and logit but really close to both, practically it has just minor difference. 

Additionally, I estimated a logit model with splines on distance because of the suggestion at the EDA as determined the breakpoint by build-in function. In this model, there is a coefficient for hotels with distance<=1.2 and and one for distance>1.2. Thus, the marginal coefficients can be interpreted as for ususal logit, but with a distinction between the distance categories.

#### Prediction
I calculated the predicted probabilities based on the 3 models and illustrated on Figure 2. The 45 degree line helps us to compare with LMP. As we can see, the LPM predictions go below zero in some cases while the logit and probit curves flatten to 0 and 1. However, around the mean the slope (marginal difference) of the 3 models are almost almost the same.

#### Summary
In this paper I used a linear probability model, a logit, a probit and an additional logit model with splines on distance to predict the probability that a hotel is highly rated explained by distance and stars.Both the distance and stars are significant in the models. The marginal differences of the models are almost identical, however logit and probit models have the adventage that they are limited at 0 and 1. Besides, based on the Bayesian information criteria, the logit model with splines is better than any of the others.

\newpage

#
#
### Appendix

```{r Read data and data mungling, message=FALSE, warning=FALSE, include=FALSE}
hotels_europe_price <- read_csv("https://osf.io/p6tyr/download")
hotels_europe_features <- read_csv("https://osf.io/utwjs/download")
# Join the 2 datatable by hotel_id
data <- left_join(hotels_europe_price, hotels_europe_features, by = "hotel_id")
# Filter for Rome
dt <- data %>%  filter(city_actual=="Rome")
rm(hotels_europe_price,hotels_europe_features,data)
# Accomodation type
dt %>% group_by(accommodation_type) %>% count()
dt<- dt %>% filter(accommodation_type=="Hotel")
dt %>% group_by(hotel_id) # Still 5327 observation and  1455 distinct hotels


dt<- dt %>% mutate( 
     highly_rated = dplyr::case_when(
      rating < 4  ~ 0,
      rating >= 4 ~ 1
    ),
    lnprice = log( price ),
    rating = as.numeric(rating)
) 

dt %>% group_by(stars) %>% count()
# Remove NA-s and not integer stars (16 observations has 3.5 stars)
dt <- dt[!is.na(dt$distance) & !is.na(dt$stars) & !is.na(dt$highly_rated) & dt$stars%%1==0, ]


summary(dt)

```

```{r Descriptive statistics, echo=FALSE}

P95 <- function(x){quantile(x,0.95,na.rm=T)}
P05 <- function(x){quantile(x,0.05,na.rm=T)}

datasummary( price + distance + offer+  weekend + holiday  + stars + rating +
            (`number of nights` = nnights ) + 
             (`Log(price)` = lnprice ) + 
             (`highly rated` = highly_rated ) ~
             Mean + Median + SD + Min + Max + P05 + P25 + P75 + P95, 
             data = dt ,
             title = 'Descriptive statistics') %>% 
      kable_styling(latex_options = c("HOLD_position","scale_down"))


```
### Correlation matrix
```{r EDA, echo=FALSE, message=FALSE, warning=FALSE}
## CORRELATION PLOT
# numerical variables to include in the correlation matrix:
# price, offer,weekend, nnights, distance, stars, rating, ratingta, distance_alter

num_table=dt[,c(2,3,7,8,9,12,13,14,25,26)]
correlation<-cor(num_table)
corrplot(correlation, addCoef.col = 'black',tl.col = "black", tl.cex = 0.9, tl.srt=45,number.cex=0.7, number.digits=2 ) 

#chart.Correlation(num_table, histogram = T, pch=19) 


```

\newpage
```{r My theme, message=FALSE, warning=FALSE, include=FALSE}
theme_gyongyver<-function(base_size=12){
  # Use the basic properties of theme_bw
  theme_light() %+replace% 
    
    # Change the items
    theme(
      # The grids on the background
      panel.grid.major  = element_line(color = "slategray2"),
      panel.grid.minor = element_line(color="grey90"),
      # The background color
      panel.background  = element_rect(fill = "grey95"),
      # the axis line
      axis.line         = element_line(color = "navyblue"),
      # Littel lines called ticks on the axis
      axis.ticks        = element_line(color = "navy"),
      # Numbers on the axis
      axis.text         = element_text(color = "navy"),
      # NEW ONES
      # rectangle element
      rect = element_rect(fill="grey10",colour = "white"),
      # axis title
      axis.title = element_text(colour="mediumblue"),
      # plot background
      plot.background = element_rect(fill="white"),
      # title
      plot.title = element_text(family = "", colour="midnightblue", size=12, hjust = 0, vjust=0.8),
      plot.subtitle = element_text(family = "", colour="midnightblue", size=10, hjust = 0),
      #caption
      plot.caption = element_text(size = 9, colour = "steelblue3", hjust = 1),
      #legend
      legend.background = element_rect(fill = "grey80", colour = "grey90"),
      legend.text = element_text(colour = "black"),
      panel.border = element_blank()
      
    )
  
}

```
#
#
### Figure 1: Exploratory data analysis
```{r EDA2, echo=FALSE, message=FALSE, warning=FALSE}
f1<-ggplot(dt)+
  geom_point(aes(x=distance, y=rating), alpha=0.5)+
  geom_smooth(aes(x=distance, y=rating), method = "loess", color="red")+
  labs(
    title = "Rating and distance",
    x="Distance",
    y="Rating"
    
  )+
  theme_gyongyver()+
  scale_x_continuous(breaks=seq(0,16,2))
  

f2<-ggplot(dt)+
  geom_point(aes(x=stars, y=rating), alpha=0.5)+
  geom_smooth(aes(x=stars, y=rating), method = "loess", color="red")+
  labs(
    title = "Rating and stars",
    x="Stars",
    y="Rating"
  )+
  theme_gyongyver()

annotate_figure(
  ggarrange(f1,f2),
  top =text_grob("Exploratory data analysis", 
                color = "midnightblue", face = "bold", size = 14)
)


```


```{r Modelling, echo=FALSE , results='asis'}
#Estimate linear probability, logit, and probit models with distance and stars as explanatory variables. 
#You may add other variables if you wish.

LPM <- feols(highly_rated ~ distance + stars, data=dt,vcov = 'hetero')

Logit <- feglm (highly_rated ~ distance + stars, family = binomial(link="logit"),vcov = 'hetero', data=dt)

Logit_marginal <- logitmfx(highly_rated ~ distance + stars, robust=TRUE, data=dt)

Probit <- feglm( highly_rated ~ distance + stars, family = binomial(link = "probit"),vcov = 'hetero', data = dt)

Probit_marginal <- probitmfx(highly_rated ~ distance + stars,robust=TRUE, data=dt)

# Splines for distance
library(segmented)
# find the optimal breakpoint
reg <- lm( highly_rated ~ distance , data = dt )
fit_dist <- segmented( reg , seg.Z = ~distance, psi = list( distance=1 ) )
#summary(fit_dist) #1.2 is the breakpoint

Logit_splines <- feglm (highly_rated ~lspline(distance,1.2)  + stars, family = binomial(link="logit"),vcov = 'hetero', data=dt)

splines_margin<-logitmfx(highly_rated ~lspline(distance,1.2)  + stars,robust=TRUE, data=dt)

#stargazer::stargazer(LPM,Logit, Probit, type = "latex", header = FALSE)
 
cm <- c('(Intercept)' = 'Constant')
msummary(list("LPM"=LPM, "Logit"=Logit,"Logit marginal"=Logit_marginal,"Probit"=Probit,"Probit marginal"=Probit_marginal,"Splines logit"=Logit_splines,"Splines marginal"=splines_margin),
         title = "Estimated models",
         fmt="%.3f",
         statistic = "std.error",
         gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|R2|PseudoR2',
         stars=c('*' = .05, '**' = .01),
         coef_rename = cm,
         notes = ''
) %>% kable_styling(latex_options = c("HOLD_position","scale_down"))


```
#
#
### Figure 2: Predicted probabilities
```{r Prediction plot, echo=FALSE, message=FALSE, warning=FALSE}
# Predicted probabilities by the models
dt$pred_lpm<-predict(LPM)
dt$pred_logit<-predict(Logit, type="response")
dt$pred_probit<-predict(Probit, type = "response")
dt$pred_spline<-predict(Logit_splines, type="response")

g1<-ggplot(dt)+
  geom_smooth(aes(x=pred_lpm, y=pred_logit),alpha=0.5 , color="blue", size=0.2, method = "loess")+
  geom_point(aes(x=pred_lpm, y=pred_logit), alpha=0.1, color="blue")+
  labs(
     title="Logit model",
     x="Predicted probability by LPM",
     subtitle = "with 45 degree line",
     y="Predicted probability by Logit model"
  )+
  theme_gyongyver()+
  geom_line(aes(x=pred_lpm, y=pred_lpm), color="red")+
  scale_x_continuous(limits = c(-0.1,0.8), breaks = seq(0,0.8,0.2))+
  scale_y_continuous(limits = c(-0.1,0.8), breaks = seq(-0.1,0.8,0.1))




g2<-ggplot(dt)+
  geom_smooth(aes(x=pred_lpm, y=pred_probit),alpha=0.5 , color="darkgreen", size=0.2, method = "loess")+
  geom_point(aes(x=pred_lpm, y=pred_probit), alpha=0.1, color="darkgreen")+
  labs(
     title="Probit model",
     subtitle = "with 45 degree line",
     x="Predicted probability by LPM",
     y="Predicted probability by Probit model"
  )+
  theme_gyongyver()+
  geom_line(aes(x=pred_lpm, y=pred_lpm), color="red")+
  scale_x_continuous(limits = c(-0.1,0.8), breaks = seq(0,0.8,0.2))+
  scale_y_continuous(limits = c(-0.1,0.8), breaks = seq(-0.1,0.8,0.1))

g3<-ggplot(dt)+
  geom_smooth(aes(x=pred_lpm, y=pred_spline),alpha=0.5, color="lightblue", size=0.2, method = "loess")+
  geom_point(aes(x=pred_lpm, y=pred_spline, color=as.factor(highly_rated)), alpha=0.5)+
  labs(
     title="Logit model with splines",
     x="Predicted probability by LPM",
     subtitle = "with 45 degree line",
     y="Predicted probability by Logit model with splines",
     color="Highly rated"
  )+
  theme_gyongyver()+
  geom_line(aes(x=pred_lpm, y=pred_lpm), color="red")+
  scale_x_continuous(limits = c(-0.1,0.8), breaks = seq(0,0.8,0.2))+
  scale_y_continuous(limits = c(-0.1,0.8), breaks = seq(-0.1,0.8,0.1))
#g3

annotate_figure(ggarrange(g1,g2),
                top =text_grob("Predicted probability that a hotel is highly rated", 
                color = "midnightblue", face = "bold", size = 14) )
```

