---
title: "Data Analysis 2 Term Project"
author: "Gyongyver Kamenar (2103380)"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: no
    toc: no
    extra_dependencies: ["float"]
header-includes:
 \usepackage{float}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	fig.pos = "H"
)
#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
# Set graph size
#knitr::opts_chunk$set(echo = FALSE, out.width = "50%" )#fig.asp = 0.5, fig.width = 7, out.width = "90%" )

#rm(list=ls())

# Libraries

library(tidyverse)
library(lspline)
library(fixest)
library(modelsummary)
library(ggpubr)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(corrplot)
library(scales)
library(bookdown)

# Get the data

#df <-read.csv('https://raw.githubusercontent.com/gyongyver-droid/ceu-data-analysis/master/Assignment3/clean_data.csv', encoding = "UTF-8")

df<-read.csv('clean_data.csv', encoding = "UTF-8")

```

## Introduction

This paper analyze the differences of English correctness between people with several different native languages. Nations have different linguistic characteristics, cultural and social norms, and education systems as well. Therefore, people with different nationalities or native languages have differences in  English learning. That's why, the motivation of this study is to see, that whether there are differences in English correctness between English learners of different native languages controlled for other relevant factors. It is a relevant question for everyone, because if there are significant differences, scientist can further investigate the main reasons behind it. It might have policy implications on education systems or recommendation of learning practices, methods and habits for English learners. 

## Data

### Origin of the data

The used dataset was origninally collected by Hartshorne et. al. (2018) who studied the critical period for second language acquisition. He collected data from 669,498 respondents through online English quizes. The ca. 10 minute-long English grammar quize was deliberately developed to enabled the researcher to measure the respondent's syntactic English knowledge and also provided several demographic variables. The critical items of the quize are the diagnostics of proficiency and there were additional items to distinguish between dialects. In my  analysis I used the percentage of correct critical items as a proxy of English correctness.

-  **English correctness** : percentage of critical quize items correct

Further information  about the data is available (here)[<https://osf.io/pyb8s/>] and in Hartshorne et. al. (2018).

### Explanatory variables

In this subsection I list the available variables in the dataset that I found relevant explanatory variable. In some cases I did not use a given variable in my analysis and I explain the reason.

-   \textbf{Age} : age of subject (numeric)
-   \textbf{Gender} : gender of the subject, either male, female or other (categorical)
-   \textbf{Native language} : subject's native language(s) (categorical)
-   \textbf{Native English} : English is the subjects native language (binary)
-   \textbf{Primary language} : subject's primary language(s) now (categorical)
-   \textbf{Primary English} : subject's primary language is English (binary)
-   \textbf{Psychiatric} : subject repoted any psychiatric disorders (binary)
-   \textbf{Starting age of English learning} : age at start of English learning (numerical)
-   \textbf{Language status} : either monoeng (native speaker of English only), bileng (native speaker of English + at least one other lang), immersion learner(spending at least 90% of their life since age of first exposure in an
English-speaking country), or non-immersion learner(spent at most 10% of post-exposure life in an English-speaking country
and no more than 1 year in total) (categorical)
-   \textbf{Years of English "learning"} : age - starting age of English learning (numerical)
-   \textbf{Education} : Highest level of education (categorical)

 Variables I found relevant but did not use:


-   \textbf{Dyslexia} : subject reported difficulty with reading (binary)

Reason: there was no variation in this variable (noone with dyslexia)

-   \textbf{Live with English} : subject lives with any native speakers (binary) 

Reason: majority of the data were NULL values indicating missing observation and I cannot be sure what it means

-   \textbf{Countries} : countries subject lived in (categorical) 

Reason: too much distinct categories (thousands) with just few subjects, moreover most subject lived in 2 or more countries

-  \textbf{Current country}: country currently lived in (categorical) 

Reason: too much distinct values (hundreds) and NAs

-  \textbf{English country years}: number of years living in English speaking countries 

Reason: most observations are missing and a similar feature is captured by Language status variable

```{r Data cleaning and mungling, message=FALSE, warning=FALSE, include=FALSE}
# Ckack the variables
summary(df)

## VARIABLES I WOULD DEFINITELY USE: 
# age, 
df %>%  group_by(age) %>% count() %>% arrange(age)
# gender, 
df %>%  group_by(gender) %>% count() %>% arrange(desc(n))
# natlangs/nat_eng, 
df %>%  group_by(natlangs) %>% count() %>% arrange(desc(n))
df %>%  group_by(nat_Eng) %>% count() %>% arrange(desc(n))
# primelangs/prime_eng, 
df %>%  group_by(primelangs) %>% count() %>% arrange(desc(n))
df %>%  group_by(prime_Eng) %>% count() %>% arrange(desc(n))
# psychiatric
df %>%  group_by(psychiatric) %>% count() %>% arrange(desc(n))
# highest level of education
df %>%  group_by(education) %>% count() %>% arrange(desc(n))
# Eng_start
df %>%  group_by(Eng_start) %>% count() %>% arrange(desc(n))
# Eng_little
df %>%  group_by(Eng_little) %>% count() %>% arrange(desc(n))
# correct / elogit
df %>%  group_by(correct) %>% count() %>% arrange(desc(n))


## Set base categories and variable type

df <-df %>% mutate(
  age = as.numeric(age),
  Eng_start = as.numeric(Eng_start),
  prime_Eng = as.numeric(prime_Eng),
  education=as.factor(education),
  natlangs = as.factor(natlangs),
  Eng_little = as.factor(Eng_little),
  learning = age - Eng_start,
  education = relevel(education,ref = "Haven't Finished High School (less than 13 years ed)"),
  natlangs = relevel(natlangs, ref = "English"),
  Eng_little = relevel(Eng_little,ref = "monoeng")
  
)


#Filter age 
df <- filter(df, df$age>=14)

# Filter error in English correctness
df %>%  group_by(correct) %>% count() %>% arrange(correct)
df <- filter(df, !(df$correct<0.5 & df$natlangs=="English"))

# Datatable with just non-english
foreign<-filter(df,df$nat_Eng==0)

```

### Scope

The data includes bilinguals and immersion learners as well, which is beyond the scope of my research question. The study's interests are subject with 1 native language (not bilingual) and in case it is not English, we want to investigate subject who are non-immersion learners (spent at most 10% of post-exposure life in an English-speaking country and no more than 1 year in total).  Thus the language status is filtered  for monoenglish or non-immersion.

Furthermore, I analyze subject with 1 native language and there are several different categories in the dataset. I decided to analyze just the top 10 native languages based on the number of subject in the dataset, namely; English, Finnish, Turkish, German, Hungarian,Russian, Dutch, Polish, Swedish and Spanish in the respective order. Even from the least frequent native language Spanish there were more then 9200 subjects in the filtered dataset.

Age is also a relevant factor in English correctness, so subjects at least 14 years old to make they are able to learn a language.

### Descriptive statistics

Please see below the descriptive statistics of the numerical variables.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Sample selection
df<- df %>% select( age, gender, education,
                     Eng_start, natlangs, nat_Eng, primelangs, prime_Eng,
                     psychiatric , correct, elogit, Eng_little,learning ) %>% drop_na()

P95 <- function(x){quantile(x,0.95,na.rm=T)}
P05 <- function(x){quantile(x,0.05,na.rm=T)}
datasummary( (`Age` = age ) +  
             (`Starting age of English learning` = Eng_start ) + 
             (`Psychiatric disorder` = psychiatric) + 
             (`English correctness` = correct) +
               (`Years of English learning` = learning) +
             (`Log(English correctness)` = elogit)  ~
             Mean + Median + SD + Min + Max + P05 + P25 + P75 + P95 , 
             data = df ,
             title = 'Descriptive statistics') %>% 
      kable_styling(latex_options = c("HOLD_position","scale_down"), font_size = 11)
```

The number of observations is `r length(!is.na(df$correct))` for all of our key variables.
The age of subjects is filtered to be at least 14 years, because below that they might not able to learn a language properly. There were some errors in the English correctness variable, because some native English people had correctness below 0.5. There is hardly and correctness score below 0.5 even for non-native English people, so these English people most probably did not make the effort to properly do the quize and that is why I removed them.

We can see, that the SD of English correctness variable is really small, the mean is `r round(mean(df$correct),2)` and the median is `r round(median(df$correct),2)` so the distribution is left skewed. See the distributions Figure \@ref(fig:Distributions) .

### Exploratory data analysis 

FIrstly, I checked the correlation between the non-categorical variables, see the correlation matrix in Figure \@ref(fig:corrplot). Native English and primary English variable are obviously highly correlated. The English starting age has strong negative correlation with native English and primary English (`r round(cor(df$nat_Eng, df$Eng_start),2)`) and also with English correctness and years of English learning (because the latter is age - English starting age). Age and years of English learning have positive correlation with English correctness. Psychiatric disorder is not really correalted with anything. In the modelling phase, I will check this matrix to decide which variable to add or not to add.

```{r corrplot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Correlation matrix", fig.height=3}

# Correlation matrix
matrix<-cor(df[,c(1,4,6,8,9,10,11,13)])
colnames(matrix)<-c("Age", "English starting age","Native English","Primary English", "Psychiatric", "English correctness",
                    "Log of English correctness","Years of English learning")
rownames(matrix)<-c("Age", "English starting age","Native English","Primary English", "Psychiatric", "English correctness",
                    "Log of English correctness","Years of English learning")

corrplot(matrix,addCoef.col = 'grey50',tl.col = "black",
         tl.cex = 0.7, tl.srt=45,number.cex=0.6, number.digits=2)

```

```{r My theme, message=FALSE, warning=FALSE, include=FALSE}
theme_gyongyver<-function(base_size=11){
  # Use the basic properties of theme_bw
  theme_light() %+replace% 
    
    # Change the items
    theme(
      # The grids on the background
      panel.grid.major  = element_line(color = "slategray2"),
      panel.grid.minor = element_line(color="grey90"),
      # The background color
      panel.background  = element_rect(fill = "grey95"),
      # the axis line
      axis.line         = element_line(color = "navyblue"),
      # Littel lines called ticks on the axis
      axis.ticks        = element_line(color = "navy"),
      # Numbers on the axis
      axis.text         = element_text(color = "navy"),
      # NEW ONES
      # rectangle element
      rect = element_rect(fill="grey10",colour = "white"),
      # axis title
      axis.title = element_text(colour="mediumblue"),
      # plot background
      plot.background = element_rect(fill="white"),
      # title
      plot.title = element_text(family = "", colour="midnightblue", size=12, hjust = 0, vjust=0.8),
      plot.subtitle = element_text(family = "", colour="midnightblue", size=10, hjust = 0),
      #caption
      plot.caption = element_text(size = 9, colour = "steelblue3", hjust = 1),
      #legend
      legend.background = element_rect(fill = "grey80", colour = "grey90"),
      legend.text = element_text(colour = "black"),
      panel.border = element_blank()
      
    )
  
}

```



```{r Distribution, warning=FALSE, include=FALSE}
# correct
p1 <- ggplot( df , aes(x = correct)) +
  geom_histogram(  fill='navyblue', color = 'white' ) +
  labs(
    title = "English correctness",
    y = 'Count',
    x = "Correct English test items (%)") +
  theme_gyongyver()+
  scale_x_continuous(limits = c(0.65,1),labels = scales::percent)+
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3))

# logit of correct
p2 <- ggplot( df , aes(x = elogit)) +
  geom_histogram(  fill='navyblue', color = 'white' ) +
  labs(
    title = "Log of English correctness",
    y = '',
    x = "Log(Correct English test items%)") +
  theme_gyongyver()+
  scale_x_continuous(limits = c(0,6),breaks=seq(0,6,1))+
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3))

# age
p3 <- ggplot( df , aes(x = age)) +
  geom_histogram(  fill='navyblue', color = 'white' ) +
  labs(
    title = "Age",
    y = 'Count',
    x = "Age") +
  theme_gyongyver()+
  scale_x_continuous(limits = c(10,75),breaks=seq(10,75,10))+
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3))

# age at start of English learning
p4 <- ggplot( df , aes(x = Eng_start)) +
  geom_histogram( fill='navyblue', color = 'white', bins = 12 ) +
  labs(
    title = "Starting age of English learning",
    y = '',
    x = "Age at start of English learning") +
  theme_gyongyver()+
  scale_x_continuous(limits = c(-5,30))+
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3))



distribution_figs <- ggarrange(p1, p2,p3, p4,
                       hjust = -0.3,
                       ncol = 2, nrow = 2)



#plot(df$natlangs, df$correct)
```

As the focus is English correctness for different native languages, the Figure \@ref(fig:Boxplot) shows the boxplot for each native language. The differences in English correctness are visible on the boxplot.

```{r Boxplot, echo=FALSE, fig.cap="Boxplot of English correctness", fig.height=3, fig.width=7, message=FALSE, warning=FALSE, fig.pos="H"}
ggplot(df,aes(x=natlangs, y=correct))+
  geom_boxplot(fill="steelblue4", color="black")+
  geom_violin(width=3, alpha=0.1, fill="skyblue2")+
  theme_gyongyver()+
  theme(axis.text.x = element_text(size = 9))+
  labs(
    title = "Boxplot of English correctness for native languages",
    x="",
    y= "English correctness"
  )+
  scale_y_continuous(labels = scales::percent)

```



The key pattern of association is:

```{r Smoothed scatterplots, message=FALSE, warning=FALSE, include=FALSE}

# Smoothed Scatterplots
f1<-ggplot(df, aes(x=Eng_start, y=correct))+
  geom_point( alpha=0.01, color="navyblue")+
  geom_smooth(method = "gam", color="red")+
  facet_wrap(~natlangs, ncol = 4)+
  theme_gyongyver()+
  labs(
    title = "Pattern of starting age of English learning",
    x = "Starting age of English learning",
    y= "English correctness"
  )
  

f2<-ggplot(df, aes(x=age, y=correct))+
  geom_point( alpha=0.01, color="navyblue")+
  geom_smooth(method = "gam", color="red")+
  facet_wrap(~natlangs, ncol = 4)+
  theme_gyongyver()+
  labs(
    title = "Pattern of age in English correctness",
    x = "Age",
    y= "English correctness"
  )

  
```



```{r Binscatter, message=FALSE, warning=FALSE, include=FALSE}

StatBinscatter <- ggplot2::ggproto(
  "StatBinscatter", 
  Stat,
  compute_group = function(data, scales, bins = 10) {
    bins     <- min(floor(nrow(data)/10), bins)
    x_bin    <- ggplot2::cut_number(data$x + 1e-12*runif(nrow(data)), bins)
    x_means  <- stats::ave(data$x, x_bin, FUN = mean)
    y_means  <- stats::ave(data$y, x_bin, FUN = mean)
    y_se     <- stats::ave(data$y, x_bin, FUN = sd)
    y_obs    <- stats::ave(data$y, x_bin, FUN = length)
    result   <- data.frame(x    = x_means, 
                           y    = y_means, 
                           ymax = y_means + 1.96*y_se/sqrt(y_obs),
                           ymin = y_means - 1.96*y_se/sqrt(y_obs))
    result   <- unique(result)
    return(result)
  },
  required_aes = c("x", "y")
)

stat_binscatter <- function(mapping = NULL, data = NULL, geom = "point",
                            position = "identity", na.rm = FALSE, show.legend = NA, 
                            inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatBinscatter, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
    )
}

bs1<-ggplot(foreign, aes(x=Eng_start, y=elogit))+
  geom_point(alpha=.01, color="navyblue")+
  stat_binscatter(color="red",bins=30,geom = "pointrange")+
  labs(
    title = "Binscatter plot",
    subtitle = "Not native English",
    x= "Starting age of English learning",
    y= "Log(English correctness)"
  )+
  theme_gyongyver()


  
```

How will you include this in your model?

Short description on the other variables: 2-10 sentence depends on the amount of variables you have. You should reference your decisions on the graphs/analysis which are located in the appendix.

## Models


```{r Models, message=FALSE, warning=FALSE, include=FALSE}

# reg1: NO control, simple linear regression
reg1 <- feols( correct ~ natlangs, data = foreign , vcov = 'hetero' )
summary(reg1)

# reg2: With controls
reg2<- feols( correct ~ age + gender + natlangs+ education +psychiatric+ Eng_start , data = df , vcov = 'hetero' )
summary(reg2)

# reg3: additional control for starting age of English learning
#   Is your parameter different? Is it a confounder?

reg3 <-feols( correct ~ age + gender + natlangs+ education +psychiatric +learning  , data = df , vcov = 'hetero' )
summary(reg3)
##
# reg4: reg3 + slines of starting age of English learning
reg4 <- feols( correct ~ age + gender + natlangs+ education + psychiatric +lspline(Eng_start,2) , data = df , vcov = 'hetero' )
summary(reg4)
#
# reg5: reg4 + salary with P.L.S, knots at 35 and 40, exptot, log of income and scratio
reg5 <- feols( correct ~ age + gender + natlangs+ education + psychiatric +lspline(Eng_start,1) ,
               data = df , vcov = 'hetero' )
reg5

etable(reg1,reg2,reg3,reg4,reg5)

# Test difference
library(car)
linearHypothesis(reg5, "natlangsPolish - natlangsHungarian = 0")
linearHypothesis(reg4,"lspline(Eng_start, 2)1 - lspline(Eng_start, 2)2 = 0")
linearHypothesis(reg5, "gendermale - genderother = 0")
# Naming the coefficients for pretty output
alpha  <- round( reg5$coeftable[1,1] , 2 )
b1 <- round( reg5$coeftable[2,1] , 2 )
b2 <- round( reg5$coeftable[3,1] , 2 )

```

My preferred model is:

score = $`r alpha`$ $`r b1`$ $( student/teacher < 18)$ $`r b2`$ $( student/teacher \geq 18) + \delta Z$

where $Z$ are standing for the controls, which includes controlling for english language, lunch, other special characteristics and wealth measures. From this model we can infer:

- when every covariates are zero, students expected to have grade score of $`r alpha`$
- when the student to teacher is one unit larger, but below the value of 18, we see students to have on average $`r abs(b1)`$ smaller grades.
- when the student to teacher is one unit larger, with the value above or equal to 18, we see students to have on average $`r abs(b2)`$ smaller grades.

However, based on the heteroskedastic robust standard errors, these results are statistically non different from zero. To show that, I have run a two-sided hypothesis test:
$$H_0:=\beta_1 - \beta_2 = 0$$
$$H_A:=\beta_1 - \beta_2 \neq 0$$
I have the t-statistic as `r round( reg5$coeftable[2,3] , 2 )` and the p-value as `r round( reg5$coeftable[2,4] , 2 )`, which confirms my conclusion.

We compare multiple models to learn about the stability of the parameters. Bla-bla:

```{r Regtable, echo=FALSE, message=FALSE, warning=FALSE}
##
# Summarize our findings:
varname_report <- c("(Intercept)" = "Intercept",
                   "natlangsDutch" = "Native Dutch",
                   "natlangsFinnish" = "Native Finnish",
                   "natlangsGerman" = "Native German",
                   "natlangsHungarian" = "Native Hungarian",
                   "natlangsPolish" = "Native Polish",
                   "natlangsRussian" = "Native Russian",
                   "natlangsSpanish" = "Native Spanish",
                   "natlangsSwedish" = "Native Swedish",
                   "natlangsTurkish" = "Native Turkish",
                   "lspline(Eng_start, 2)1" = "Starting age of English learning (<2)",
                   "lspline(Eng_start, 2)2" = "Starting age of English learning (>=2)",
                   "age" = "Age",
                   "gendermale" = "Male",
                   "genderother" = "Other gender",
                   "educationHigh School Degree (12-13 years)" = "HighSchool Degree (12-13years ed)",
                   "educationSome Undergrad (higher ed)" = "Some Undergrad (highered)",
                   "educationUndergraduate Degree (3-5 years higher ed)" = "Undergraduate Degree (3-5years highered)",
                   "educationSome Graduate School" = "Some Graduate School",
                   "educationGraduate Degree" = "Graduate Degree",
                   "Eng_start" = "Starting age of English learning",
                   "psychiatric" = "Psychiatric",
                   "learning" = "Years of English learning",
                   "Eng_littlelittle" = "Non-immersion learner"
                   
                   )


# Note: coefstat = 'confint' is just an example, usually you need to report se.
style_noHeaders = style.tex(var.title = "", fixef.title = "", stats.title = " ")


# kable( etable( reg1 , reg2 , reg3 , reg4 , reg5 ,
#         title = 'Average test scores for 4th graders',
#         dict = varname_report,
#         drop = vars_omit ,
#         group = groupConf ,
#         se.below = T,
#         coefstat = 'se',
#         fitstat = c('n','r2'),
#         se.row = F,
#         depvar = F ) , 
#         col.names = c('(1)','(2)','(3)','(4)','(5)'),
#        "latex", booktabs = TRUE,  position = "H",
#        caption = 'Models to uncover relation between test score and student to teacher ratio') %>% kable_styling(latex_options = c("hold_position","scale_down"))

msummary(list("Simple"=reg1, "Multiple"=reg2,"Multiple with starting age"=reg3,
              "Splines"=reg4,"Splines"=reg5),
         title = "Estimated models",
         fmt="%.3f",
         statistic = "std.error",
         gof_omit = 'DF|Deviance|F|R2 Adj.|AIC|PseudoR2',
         stars=c('**' = .05, '***' = .01),
         coef_rename = varname_report,
         notes = ''
) %>% kable_styling(latex_options = c("HOLD_position","scale_down"),
                    font_size = 11)


```





## Robustness check / 'Heterogeneity analysis'

Task: calculate and report t-tests for each countries. 



## Conclusion

HERE COMES WHAT WE HAVE LEARNED AND WHAT WOULD STRENGHTEN AND WEAKEN OUR ANALYSIS.

\newpage 
## Appendix

Here comes all the results which are referenced and not essential for understanding the MAIN results.

```{r Distributions, echo=FALSE, fig.cap="Distributions", message=FALSE, warning=FALSE, out.height='100%', out.width='100%', fig.pos="H"}
distribution_figs
```

```{r echo=FALSE, fig.cap="Frequency of native languages", message=FALSE, warning=FALSE}
# frequency of native languages
df %>%  group_by(natlangs) %>% count() %>% arrange(desc(n)) %>% 
  ggplot(aes(x=natlangs, y=n))+
  geom_bar(stat='identity', fill="navyblue")+
  geom_text(aes(label=n), vjust=-0.3, size=4) +
  theme_gyongyver()+
  labs(
    title = "Frequency of native languages",
    x="",
    y=""
  )+
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3))+
  theme(axis.text.x = element_text(size = 9))

```

```{r echo=FALSE, fig.cap="Binned, message=TRUE, warning=TRUE}

ggarrange(bs1)

```

```{r echo=FALSE, fig.cap="Smoothed pattern", message=FALSE, warning=FALSE}
# Smoothed pattern in appendix
f1

f2

```


\newpage
## Bibliography

Hartshorne, J. K., Tenenbaum, J. B., & Pinker, S. (2018). A critical period for second language acquisition: Evidence from 2/3 million English speakers. Cognition, 177, 263-277.
